---
title: "Binding density coverage tracks"
subtitle: "the final version"
output: html_notebook
---

```{r, warning=FALSE}
library(tibble)
library(purrr)
library(ggpubr)
library(ggsignif)

source("~/HexamerBias/artificial_coverage/compare_peaks.r")
```

## C.elegans WGS
```{r}
predicted.bw <- '~/mnt/edann/VAN2423_onePreamp/cov_prediction/WBcel235.random.42.srt.noOvs.noChrcele_pt_noBS_predcoverage.artCov.bw'
experimental.bw <- '~/mnt/edann/VAN2423_onePreamp/CG-pbat-gDNA-CeleTotal-noBS-1preAmp-handMix_lmerged_R1.srt.dedup.bw'
cele.new <- make.predVSexp.track(predicted.bw = predicted.bw,
                                 experimental.bw = experimental.bw)

```
Saving example tracks
```{r}
for (n in 1:10) {
  smp <- sample(cele.new,1)[[1]]
  sp <- sp.cele.new$spear[smp$range.id[1]]
  nice.plotTrack(smp) +
    annotate('text',x = smp@ranges@start[500], y=max(smp$score), label=paste("Spearman's rho =", round(sp, 2)), size=8) +
    ggsave(paste0("~/AvOwork/output/MyData/coverage_tracks/cele_noBS_smp", n, '.pdf'), width=10)
}
```


#### Spearman correlation between exp and predicted
```{r}
sp.cele.new<- compare.spear.real(cele.new,name.real = 'CElegans_VAN2423')
pcc.cele.new<- compare.spear.real(cele.new,name.real = 'CElegans_VAN2423', method = 'pearson')
sp.cele.new$spear %>% median

sp.cele.new$p
ggsave("~/AvOwork/output/MyData/coverage_tracks/spearman_dist_vs_random.pdf")
pcc.cele.new$p
```
```{r}
high.spear <- cele.new[which(sp.cele.new$spear>0.8)]

for (n in 1:10) {
  smp <- sample(high.spear,1)[[1]]
  sp <- sp.cele.new$spear[smp$range.id[1]]
  nice.plotTrack(smp) +
    annotate('text',x = smp@ranges@start[500], y=max(smp$score), label=paste("Spearman's rho =", round(sp, 2)), size=8) +
    ggsave(paste0("~/AvOwork/output/MyData/coverage_tracks/cele_noBS_highspear", n, '.pdf'), width=10)
}

```


#### Coverage dsitribution per spearman correlation
```{r}
mean.norm.cov <- sapply(cele.new, function(x) mean(x$score))
mean.df <- data.frame(mean.cov = mean.norm.cov, id=names(mean.norm.cov))

cov.cor.df <- as.data.frame(sp.cele.new$spear) %>%
  rownames_to_column(var='id') %>%
  inner_join(., mean.df) %>%
  # mutate(spear.group = cut(sp.cele.new$spear, breaks = seq(-1,1,by = 0.25), include.lowest = TRUE)) 
  mutate(spear.group = cut(sp.cele.new$spear, breaks = c(-1,seq(0,1,by = 0.25)), include.lowest = TRUE)) 

pval.df <- compare_means(mean.cov ~ spear.group, data=cov.cor.df, method='t.test') %>%
  mutate(y.pos=6e-07)
ggplot(cov.cor.df, aes(y=mean.cov, x=spear.group)) +
  geom_boxplot(varwidth = TRUE, 
               outlier.shape = NA, 
               fill=NA) +
  coord_cartesian(ylim=c(-1,2)) +
  theme_classic() +
  xlab('Spearman corr. coefficient') + ylab('Mean normalized coverage') +
  theme(axis.title = element_text(size = 20), axis.text = element_text(size=10), title = element_text(size=22)) +
  geom_signif(comparisons = list(levels(cov.cor.clean$spear.group)[c(2,5)],
                                  levels(cov.cor.clean$spear.group)[c(2,4)], 
                                 levels(cov.cor.clean$spear.group)[3:2],
                                 levels(cov.cor.clean$spear.group)[c(3,4)]
                                 ), 
              test='t.test',
              map_signif_level = TRUE, y_position = seq(1.1, 1.7, length.out = 4), tip_length = 0.001
              ) 
```

##### Try taking smaller regions (1000 bps)

```{r}
cele.new.1kb <- make.predVSexp.track(predicted.bw = predicted.bw,
                                 experimental.bw = experimental.bw,
                                 reg.length = 1000)

```
```{r}
spear.cele.1kb <- compare.spear.real(cele.new.1kb, name.real = 'cele_WGS_1kb')
spear.cele.1kb$p
```
```{r}
smp <- sample(cele.new.1kb,1)[[1]]
spear.cele.1kb$spear[smp$range.id[1]]
nice.plotTrack(smp)
```

## C.elegans BS-seq
```{r}
predicted.bw <- '~/mnt/edann/VAN2423_onePreamp/cov_prediction/WBcel235.random.42.srt.noOvs.noChrCG-pbat-gDNA-CeleTotal-withBS-1preAmp-handMix_lmerged_R1_trimmed_bismark_bt2.predcoverage.artCov.bw'
experimental.bw <- '~/mnt/edann/VAN2423_onePreamp/CG-pbat-gDNA-CeleTotal-withBS-1preAmp-handMix_lmerged_R1_trimmed_bismark_bt2.deduplicated.srt.bw'
cele.bs <- make.predVSexp.track(predicted.bw = predicted.bw,
                                 experimental.bw = experimental.bw)

```
```{r}
nice.plotTrack(sample(cele.bs,1)[[1]])
```
#### Spearman correlation between exp and predicted
```{r}
sp.cele.bs <- compare.spear.real(cele.bs,name.real = 'CElegans_VAN2423_BSseq')
sp.cele.bs$p + ggsave("~/AvOwork/output/MyData/coverage_tracks/spearman_dist_vs_random_BS.pdf")
```

#### Coverage dsitribution per spearman correlation
```{r}
mean.norm.cov <- sapply(cele.bs, function(x) mean(x$score))
mean.df <- data.frame(mean.cov = mean.norm.cov, id=names(mean.norm.cov))

cov.cor.df <- as.data.frame(sp.cele.bs$spear) %>%
  rownames_to_column(var='id') %>%
  inner_join(., mean.df) %>%
  # mutate(spear.group = cut(sp.cele.bs$spear, breaks = seq(-1,1,by = 0.25), include.lowest = TRUE)) 
  mutate(spear.group = cut(sp.cele.bs$spear, breaks = c(-1,seq(0,0.5,by = 0.25),1), include.lowest = TRUE)) 

pval.df <- compare_means(mean.cov ~ spear.group, data=cov.cor.df, method='t.test') %>%
  mutate(y.pos=6e-07)
ggplot(filter(cov.cor.df, !is.na(spear.group)), aes(y=mean.cov, x=spear.group)) +
  geom_boxplot(varwidth = TRUE, 
               outlier.alpha = 0.3, 
               fill=NA) +
  coord_cartesian(ylim=c(-2,3.5)) +
  theme_classic() +
  xlab('Spearman corr. coefficient') + ylab('Mean normalized coverage') +
  theme(axis.title = element_text(size = 20), axis.text = element_text(size=10), title = element_text(size=22)) +
  geom_signif(comparisons = list(# levels(cov.cor.clean$spear.group)[c(2,5)],
                                 levels(cov.cor.df$spear.group)[c(1,2)],
                                 levels(cov.cor.df$spear.group)[c(1,3)],
                                 levels(cov.cor.df$spear.group)[c(1,4)],
                                  levels(cov.cor.df$spear.group)[c(2,4)],
                                  levels(cov.cor.df$spear.group)[c(3,4)]
                                 # levels(cov.cor.clean$spear.group)[3:2]
                                 # levels(cov.cor.clean$spear.group)[c(3,4)]
                                 ), 
              test='t.test',
              map_signif_level = TRUE, y_position = seq(2.5, 3.4, length.out = 5), tip_length = 0.001
              ) 
```

### Mouse BS
```{r}
predicted.bw <- '~/mnt/edann/VAN2423_onePreamp/cov_prediction/WBcel235.random.42.srt.noOvs.noChrCG-pbat-gDNA-CeleTotal-withBS-1preAmp-handMix_lmerged_R1_trimmed_bismark_bt2.predcoverage.artCov.bw'
experimental.bw <- '~/mnt/edann/VAN2423_onePreamp/CG-pbat-gDNA-CeleTotal-withBS-1preAmp-handMix_lmerged_R1_trimmed_bismark_bt2.deduplicated.srt.bw'
cele.bs <- make.predVSexp.track(predicted.bw = predicted.bw,
                                 experimental.bw = experimental.bw)

```
```{r}
nice.plotTrack(sample(cele.bs,1)[[1]])
```



Human dataset 
```{r}
load('~/AvOwork/rdata/human_noBS.covprofiles.RData', verbose = T)

track.human.noBS <- smooth.coverage(human_noBS.common.bw) %>%
  normalize.coverage() %>%
  add.id.2()
track.human.noBS <- lapply(split.tracks(track.human.noBS), trim.edges)
```
```{r}
spear.human <- map_dbl(track.human.noBS, corr.real)
rand.human <- randomize(track.human.noBS)

data.frame(sample='human_WGS', spearman.rho=spear.human) %>%
  bind_rows(., data.frame(sample='random', spearman.rho=rand.human)) %>%
  ggplot(., aes(sample, spearman.rho)) +
  geom_boxplot(outlier.alpha = 0.2) +
  theme_classic() +
  ggsignif::geom_signif(comparisons = list(c("human_WGS", "random")), 
              map_signif_level=TRUE) +
  theme(axis.title=element_text(size=20), 
        axis.text=element_text(size=16), 
        strip.text=element_text(size=20), 
        legend.text = element_text(size=20),
        legend.title = element_text(size=22)) +
  xlab('')

```
```{r}
mean.norm.cov <- sapply(track.human.noBS, function(x) mean(x$score))
mean.df <- data.frame(mean.cov = mean.norm.cov, id=names(mean.norm.cov))

cov.cor.df <- as.data.frame(spear.human) %>%
  rownames_to_column(var='id') %>%
  inner_join(., mean.df) %>%
  mutate(spear.group = cut(spear.human, breaks = seq(-1,1,by = 0.25), include.lowest = TRUE)) %>% 
  mutate(spear.group = cut(spear.human, breaks = c(-1,seq(0,1,by = 0.25)), include.lowest = TRUE)) 

pval.df <- compare_means(mean.cov ~ spear.group, data=cov.cor.df, method='t.test') 

ggplot(cov.cor.df, aes(y=mean.cov, x=spear.group)) +
  geom_boxplot(varwidth = TRUE, 
               outlier.shape = NA, 
               fill=NA) +
  coord_cartesian(ylim=c(-1,2)) +
  theme_classic() +
  xlab('Spearman corr. coefficient') + ylab('Mean normalized coverage') +
  theme(axis.title = element_text(size = 20), axis.text = element_text(size=10), title = element_text(size=22)) +
  # geom_signif(comparisons = list(levels(cov.cor.clean$spear.group)[c(2,5)],
  #                                 levels(cov.cor.clean$spear.group)[c(2,4)], 
  #                                levels(cov.cor.clean$spear.group)[3:2],
  #                                levels(cov.cor.clean$spear.group)[c(3,4)]
  #                                ), 
  #             test='t.test',
  #             map_signif_level = TRUE, y_position = seq(1.1, 1.7, length.out = 4), tip_length = 0.001
  #             )
  NULL
```


```{r}
low.spear <- names(sort(spear, decreasing = T))

nice.plotTrack(track.list.noBS[low.spear[2]][[1]])

```

### Mouse BS data 
Load data
```{r}
predicted.bw <- '~/mnt/edann/hexamers/strand_specific/artificial_coverage/highcov.random.42.artCov.bw'
experimental.bw <- '~/mnt/edann/hexamers/strand_specific/VAN1667_se.highcov42.bw'
VAN1667.bs <- make.predVSexp.track(predicted.bw = predicted.bw,
                                experimental.bw = experimental.bw,
                               reg.length = 5000)
```
#### Spearman correlation between exp and predicted
```{r}
sp.mm.new<- compare.spear.real(VAN1667.bs, name.real = 'MMus_VAN1667')
sp.mm.new$spear %>% median

sp.mm.new$p +
  ggsave("~/AvOwork/formatted_figs/spearman_dist_vs_random_VAN1667.pdf")
pcc.cele.new$p
```

Saving example tracks
```{r}

for (n in 1:10) {
  smp <- sample(VAN1667.bs[which(sp.mm.new$spear>0.8)],1)[[1]]
  sp <- sp.mm.new$spear[smp$range.id[1]]
  nice.plotTrack(smp) +
    annotate('text',x = smp@ranges@start[500], y=max(smp$score), label=paste("Spearman's rho =", round(sp, 2)), size=8) +
    ggsave(paste0("~/AvOwork/output/artificial_coverage/VAN1667_smp_veryhighspear_", n, '.pdf'), width=10)
}
```



#### Coverage dsitribution per spearman correlation
```{r}
mean.norm.cov <- sapply(VAN1667.bs, function(x) mean(x$score))
mean.df <- data.frame(mean.cov = mean.norm.cov, id=names(mean.norm.cov))

cov.cor.df <- as.data.frame(sp.mm.new$spear) %>%
  rownames_to_column(var='id') %>%
  inner_join(., mean.df) %>%
  # mutate(spear.group = cut(sp.mm.new$spear, breaks = seq(-1,1,by = 0.25), include.lowest = TRUE)) 
  mutate(spear.group = cut(sp.mm.new$spear, breaks = c(-1,seq(0.25,0.75,by = 0.25),1), include.lowest = TRUE)) 

ggplot(filter(cov.cor.df, !is.na(spear.group)), aes(y=mean.cov, x=spear.group)) +
  geom_boxplot(varwidth = TRUE, 
               outlier.alpha = 0, 
               fill=NA) +
  coord_cartesian(ylim=c(-0.8,2)) +
  theme_classic() +
  xlab('Spearman corr. coefficient') + ylab('Mean normalized coverage') +
  theme(axis.title = element_text(size = 20), axis.text = element_text(size=10), title = element_text(size=22)) +
  geom_signif(comparisons = list(# levels(cov.cor.clean$spear.group)[c(2,5)],
                                 levels(cov.cor.df$spear.group)[c(1,2)],
                                 levels(cov.cor.df$spear.group)[c(1,3)],
                                 levels(cov.cor.df$spear.group)[c(2,3)],
                                levels(cov.cor.df$spear.group)[c(2,4)],
                                levels(cov.cor.df$spear.group)[c(3,4)]
                                 # levels(cov.cor.clean$spear.group)[c(1,4)]
                                 # levels(cov.cor.clean$spear.group)[c(3,4)]
                                 ), 
              test='t.test',
              map_signif_level = TRUE, y_position = seq(1.1, 1.8, length.out = 5), 
              tip_length = 0.001,
              vjust=0.2
              ) +
  ggsave('~/AvOwork/formatted_figs/outliers_BS_spear.pdf')
```
