---
title: "Optimizing coverage at cpg sites"
output: html_notebook
---

```{r, warning=FALSE}
library(rtracklayer)
library(purrr)
library(Gviz)
library(zoo)
library(flux)
source("~/HexamerBias/deltaGprediction/binding_model_functions.r")
source("~/HexamerBias/artificial_coverage/compare_peaks.r")
```


Load predicted coverage track 
```{r}
cpg.bed.file <- '~/mnt/edann/pred_coverage_primer_batch_D3R/evenNreads/random.42.mm10.noOvsD3R_dens_0.25_0.25_0.25_0.25evenNreads_predcoverage.artCov.cpg_coverage.bed'
random.bed.file <- '~/mnt/edann/pred_coverage_primer_batch_D3R/evenNreads/random.42.mm10.noOvsD3R_dens_0.25_0.25_0.25_0.25evenNreads_predcoverage.artCov.random.bed'

cpg.bed.even <- import(cpg.bed.file, format = 'BED')
random.bed.even <- import(random.bed.file, format = 'BED')

# cpg.norm.bed <- normalize.coverage(cpg.bed)
# random.norm.bed <- normalize.coverage(random.bed)
boxplot(random.bed$score, cpg.bed$score, outline=F)
t.test(cpg.bed$score, random.bed$score)
```


```{r}
cpg.bed.file <- '~/mnt/edann/pred_coverage_primer_batch_D3R/evenNreads/random.42.mm10.noOvsD3R_dens_0_0.25_0_0.75evenNreads_predcoverage.artCov.cpg_coverage.bed'
random.bed.file <- '~/mnt/edann/pred_coverage_primer_batch_D3R/evenNreads/random.42.mm10.noOvsD3R_dens_0_0.25_0_0.75evenNreads_predcoverage.artCov.random.bed'

cpg.bed.best <- import(cpg.bed.file, format = 'BED')
random.bed.best <- import(random.bed.file, format = 'BED')

# cpg.norm.bed <- normalize.coverage(cpg.bed)
# random.norm.bed <- normalize.coverage(random.bed)
df <- as.data.frame(list(best.random=random.bed.best$score, best.cpgs=cpg.bed.best$score, even.cpgs=cpg.bed.even$score, even.random=random.bed.even$score)) %>%
  melt(value.name='cov.density') %>%
  mutate(batch=gsub(pattern = '\\..+', replacement = '', variable),
         region=gsub(pattern = '.+\\.', replacement = '', variable),
         ) %>%
  group_by(batch) %>%
  mutate(zscore=(cov.density-mean(cov.density, na.rm=T))/sd(cov.density, na.rm=T)) %>%
  ungroup() 

df %>%
  ggplot(., aes(y=zscore, x=batch, fill=region)) +
    geom_boxplot(outlier.alpha=0.2) 

data_frame(random.bed.best$score)
boxplot(random.bed.best$score, cpg.bed.best$score,random.bed$score, cpg.bed$score, outline=F)
t.test(cpg.bed.best$score, random.bed.best$score)
```

Try for CpG islands (more CG rich regions)
```{r}
cgi.bed.file <- '~/mnt/edann/pred_coverage_primer_batch_D3R/evenNreads/random.42.mm10.noOvsD3R_dens_0_0.25_0_0.75evenNreads_predcoverage.artCov.zscore.cgi_coverage.bed'
random.bed.file <- '~/mnt/edann/pred_coverage_primer_batch_D3R/evenNreads/random.42.mm10.noOvsD3R_dens_0_0.25_0_0.75evenNreads_predcoverage.artCov.zscore.randomCGI.bed'

cgi.bed.best <- import(cgi.bed.file, format = 'BED')
random.bed.best <- import(random.bed.file, format = 'BED')


cgi.bed.file <-  '~/mnt/edann/pred_coverage_primer_batch_D3R/evenNreads/random.42.mm10.noOvsD3R_dens_0.25_0.25_0.25_0.25evenNreads_predcoverage.artCov.zscore.cgi_coverage.bed'
random.bed.file <-  '~/mnt/edann/pred_coverage_primer_batch_D3R/evenNreads/random.42.mm10.noOvsD3R_dens_0.25_0.25_0.25_0.25evenNreads_predcoverage.artCov.zscore.randomCGI.bed'

cgi.bed.even <- import(cgi.bed.file, format = 'BED')
random.bed.even <- import(random.bed.file, format = 'BED')


bed.list <- list(best.random=random.bed.best$score, best.cgis=cgi.bed.best$score, even.cgis=cgi.bed.even$score, even.random=random.bed.even$score) 
  
df2 <- map(seq_along(bed.list), function(i) data.frame(cov.density=bed.list[[i]], variable=names(bed.list)[i])) %>%
  reduce(bind_rows) %>%
  # melt(value.name='cov.density') %>%
  mutate(batch=gsub(pattern = '\\..+', replacement = '', variable),
         region=gsub(pattern = '.+\\.', replacement = '', variable),
         ) 
  group_by(batch) %>%
  mutate(zscore=(cov.density-mean(cov.density))/sd(cov.density)) %>%
  ungroup() 


df2 %>%
  # filter(batch=='best') %>%
  # group_by(variable) %>%
  # summarise(cov=mean(cov.density), sd=sd(cov.density), batch=first(batch), region=first(region)) %>%
  ggplot(., aes(y=cov.density, x=batch, fill=region)) +
    geom_boxplot(outlier.alpha=0.01, varwidth = F) +
  ylim(-5,20)
```

Try for TSS
```{r}

compare.best.even <- function(reg.best.bed.file, random.best.bed.file, reg.even.bed.file, random.even.bed.file, plot=T, reg.name='reg'){
  cgi.bed.best <- import(reg.best.bed.file, format = 'BED')
  random.bed.best <- import(random.best.bed.file, format = 'BED')
  cgi.bed.even <- import(reg.even.bed.file, format = 'BED')
  random.bed.even <- import(random.even.bed.file, format = 'BED')
  bed.list <- list(best.random=random.bed.best$score, best.reg=cgi.bed.best$score, even.reg=cgi.bed.even$score, even.random=random.bed.even$score) 
  df2 <- map(seq_along(bed.list), function(i) data.frame(cov.density=bed.list[[i]], variable=names(bed.list)[i])) %>%
    reduce(bind_rows) %>%
    mutate(batch=gsub(pattern = '\\..+', replacement = '', variable),
           region=gsub(pattern = '.+\\.', replacement = '', variable)
           ) %>%
    mutate(region=ifelse(region=='reg', reg.name, region))
  if (plot) {
    p <- df2 %>%
    ggplot(., aes(y=cov.density, x=batch, fill=region)) +
      geom_boxplot(outlier.alpha=0.01, varwidth = F) +
      ylab("Cov. density (z-score)") +
      scale_fill_brewer(palette='YlGnBu') +
      theme_classic() +
      theme(axis.title=element_text(size=20), 
        axis.text=element_text(size=16), 
        strip.text=element_text(size=20), 
        legend.text = element_text(size=20),
        legend.title = element_text(size=22))
    p
    }
  return(p)
  }

p.ctcf <- compare.best.even(
  reg.best.bed.file = '~/mnt/edann/pred_coverage_primer_batch_D3R/evenNreads/random.42.mm10.noOvsD3R_dens_0_0.25_0_0.75evenNreads_predcoverage.artCov.zscore.ctcf_coverage.bed',
  random.best.bed.file = '~/mnt/edann/pred_coverage_primer_batch_D3R/evenNreads/random.42.mm10.noOvsD3R_dens_0_0.25_0_0.75evenNreads_predcoverage.artCov.zscore.randomCTCF.bed',
  reg.even.bed.file = "~/mnt/edann/pred_coverage_primer_batch_D3R/evenNreads/random.42.mm10.noOvsD3R_dens_0.25_0.25_0.25_0.25evenNreads_predcoverage.artCov.zscore.ctcf_coverage.bed",
  random.even.bed.file = '~/mnt/edann/pred_coverage_primer_batch_D3R/evenNreads/random.42.mm10.noOvsD3R_dens_0.25_0.25_0.25_0.25evenNreads_predcoverage.artCov.zscore.randomTSS.bed',
  reg.name = 'CTCF')


p.cgi <- compare.best.even(
  reg.best.bed.file = '~/mnt/edann/pred_coverage_primer_batch_D3R/evenNreads/random.42.mm10.noOvsD3R_dens_0_0.25_0_0.75evenNreads_predcoverage.artCov.zscore.cgi_coverage.bed',
  random.best.bed.file = '~/mnt/edann/pred_coverage_primer_batch_D3R/evenNreads/random.42.mm10.noOvsD3R_dens_0_0.25_0_0.75evenNreads_predcoverage.artCov.zscore.randomCGI.bed',
  reg.even.bed.file = "~/mnt/edann/pred_coverage_primer_batch_D3R/evenNreads/random.42.mm10.noOvsD3R_dens_0.25_0.25_0.25_0.25evenNreads_predcoverage.artCov.zscore.cgi_coverage.bed",
  random.even.bed.file = '~/mnt/edann/pred_coverage_primer_batch_D3R/evenNreads/random.42.mm10.noOvsD3R_dens_0.25_0.25_0.25_0.25evenNreads_predcoverage.artCov.zscore.randomCGI.bed',
  reg.name='CpG islands')


p.ctcf +
  ylim(-3,4) +
  ggsave('~/AvOwork/output/DE_optimization/bestFC_vs_even_ctcf_boxplot.pdf')

p.cgi +
  ylim(-5,25) +
  ggsave('~/AvOwork/output/DE_optimization/bestFC_vs_even_cgis_boxplot.pdf')
```

```{r}
p.cpgs <- compare.best.even(
  reg.best.bed.file = '~/mnt/edann/pred_coverage_primer_batch_D3R/evenNreads/random.42.mm10.noOvsD3R_dens_0_0.25_0_0.75evenNreads_predcoverage.artCov.zscore.cpg_coverage.bed',
  random.best.bed.file = '~/mnt/edann/pred_coverage_primer_batch_D3R/evenNreads/random.42.mm10.noOvsD3R_dens_0_0.25_0_0.75evenNreads_predcoverage.artCov.zscore.randomCPG.bed',
  reg.even.bed.file = "~/mnt/edann/pred_coverage_primer_batch_D3R/evenNreads/random.42.mm10.noOvsD3R_dens_0.25_0.25_0.25_0.25evenNreads_predcoverage.artCov.zscore.cpg_coverage.bed",
  random.even.bed.file = '~/mnt/edann/pred_coverage_primer_batch_D3R/evenNreads/random.42.mm10.noOvsD3R_dens_0.25_0.25_0.25_0.25evenNreads_predcoverage.artCov.zscore.randomCPG.bed',
  reg.name='CpG sites')

p.cpgs +
  coord_cartesian(ylim=c(-3,4)) 
```

Merged CpG islands
```{r}
p.cgi.merged <- compare.best.even(
  reg.best.bed.file = '~/mnt/edann/pred_coverage_primer_batch_D3R/evenNreads/random.42.mm10.noOvsD3R_dens_0_0.25_0_0.75evenNreads_predcoverage.artCov.zscore.merged.cgi_coverage.bed',
  random.best.bed.file = '~/mnt/edann/pred_coverage_primer_batch_D3R/evenNreads/random.42.mm10.noOvsD3R_dens_0_0.25_0_0.75evenNreads_predcoverage.artCov.zscore.merged.randomCGI.bed',
  reg.even.bed.file = "~/mnt/edann/pred_coverage_primer_batch_D3R/evenNreads/random.42.mm10.noOvsD3R_dens_0.25_0.25_0.25_0.25evenNreads_predcoverage.artCov.zscore.merged.cgi_coverage.bed",
  random.even.bed.file = '~/mnt/edann/pred_coverage_primer_batch_D3R/evenNreads/random.42.mm10.noOvsD3R_dens_0.25_0.25_0.25_0.25evenNreads_predcoverage.artCov.zscore.merged.randomCGI.bed',
  reg.name='CpG islands')

p.cgi.merged +
  ggsave('~/AvOwork/output/DE_optimization/bestFC_vs_even_cgis_boxplot.pdf')
```

### Predicted coverage tracks for even VS best

```{r}
best.batch <- "~/mnt/edann/pred_coverage_primer_batch_D3R/evenNreads/random.42.mm10.noOvsD3R_dens_0_0.25_0_0.75evenNreads_predcoverage.artCov.bw"
even.batch <- "~/mnt/edann/pred_coverage_primer_batch_D3R/evenNreads/random.42.mm10.noOvsD3R_dens_0.25_0.25_0.25_0.25evenNreads_predcoverage.artCov.bw"

best.even.track <- load.expVSpred.coverage(best.batch, even.batch)
track <- make.predVSexp.range(pred.bw = best.even.track$pred, exp.bw = best.even.track$exp, pred.name = 'best', exp.name = 'even') %>%
    normalize.coverage() %>%
    add.id.2(reg.length = 2000)
track.list <- lapply(split.tracks(track, reg.length = 2000), trim.edges)

```


```{r}
smp <- sample(track.list,1)[[1]]
n <- as.numeric(smp$range.id)[1]
nice.plotTrack(smp, labels = c('even', 'best')) +
  ggsave(paste0('~/AvOwork/output/DE_optimization/bestVSeven_cov_tracks/bestVSeven_cov_track_', n, '.pdf'))

```

Load track of regions of interest
```{r}
cgis.track <- import('~/mnt/edann/hexamers/annotations_bed/CpGislands_srt_mm10.bed', format = 'BED')

int.cgis <- findOverlaps(query = track, subject = cgis.track)
auc.best.roi <- compute.auc(track[queryHits(int.cgis)]$best)
auc.even.roi <- compute.auc(track[queryHits(int.cgis)]$even)

random.out <- track[-queryHits(int.cgis)]
compute.auc(random.out$even)
compute.auc(random.out$best)
```

I will try to scale everything up to not have to deal with negative values
```{r}
scaled.track <- track
score.cols <- colnames(values(scaled.track)[sapply(values(scaled.track), is.numeric)])
min.score <- min(as.matrix(scaled.track@elementMetadata[score.cols]))
for (col in score.cols) {
    scaled.track@elementMetadata[col][[1]] <- scaled.track@elementMetadata[col][[1]] + abs(min.score)
  }

coverage.yield <- function(scaled.track, roi.track){
  int.roi <- findOverlaps(query = scaled.track, subject = roi.track)
  auc.best.roi <- auc(seq_along(scaled.track[queryHits(int.roi)]$best), scaled.track[queryHits(int.roi)]$best)
  auc.even.roi <- auc(seq_along(scaled.track[queryHits(int.roi)]$even), scaled.track[queryHits(int.roi)]$even)
  random.out <- scaled.track[-queryHits(int.roi)]
  auc.even.out <- auc(seq_along(sample(random.out$even, length(queryHits(int.roi)))), sample(random.out$even, length(queryHits(int.roi))))
  auc.best.out <- auc(seq_along(sample(random.out$best, length(queryHits(int.roi)))), sample(random.out$best, length(queryHits(int.roi))))
  yield.best <- auc.best.roi/auc.best.out
  yield.even <- auc.even.roi/auc.even.out
  return(yield.best-yield.even)
  }

coverage.yield(scaled.track = scaled.track, cgis.track)
yield.best-yield.even
```

Yahooo!

##### Is this significant?
Made script `pval_yield_iteration.r` to compute the $\Delta(yield)$ for random permutations of the original even track, to test significance of the result.
```{r}
scaled.track.smp <- readRDS('~/mnt/edann/pred_coverage_primer_batch_D3R/evenNreads/coverage_yield/smp1.RDS')
randm <- scan('~/mnt/edann/pred_coverage_primer_batch_D3R/evenNreads/coverage_yield/yield_pval.smp1.CpGislands_srt_mm10.txt')
yield <- coverage.yield(scaled.track.smp, cgis.track)

data.frame(randm) %>%
  ggplot(., aes(randm)) +
  geom_histogram() +
  geom_vline(xintercept = yield, color='red')
  
(randm, breaks = 30)
```

### Computing yield for all available tracks
Used `compute_yield_track_dir.r` to compute yield for all available tracks
```{r}
cgi.yield <- read.csv('~/mnt/edann/pred_coverage_primer_batch_D3R/evenNreads/yield_CpGislands_srt_mm10.csv')

cgi.yield %>%
  mutate(track=gsub(track, pattern = 'random.+D3R_|evenN.+', replacement = '')) %>%
  mutate(delta_yield=yield-1.3973866) %>%
  mutate(fc.score=fc.scores.cgi[match(track, names(fc.scores.cgi))]) %>%
  ggplot(., aes(x=track, fill=fc.score, yield)) +
  geom_bar(stat='identity') +
  coord_flip()

```

