---
title: "Coverage fraction prediction"
output: html_notebook
---

```{r, eval=F}
source("~/HexamerBias/deltaGprediction/binding_model_functions.r")
# source('~/HexamerBias/rscripts/hexamer_sequence_functions.r')
```

```{r}
load("~/AvOwork/rdata/human_pt_all.RData")
load("~/AvOwork/rdata/cele_pt_all.RData")
load("~/AvOwork/rdata/mean_keqs_noBS_unfilt.RData")
load("~/AvOwork/rdata/D3R_pt.RData")
```

```{r}
human.keqs.df <- inner_join(avg.keqs.all, human.all.df, by=c("primer", 'template', 'dG')) 
pred.cov.human <- predict.coverage(human.keqs.df, eps=eps.human)  

cele.keqs.df <- inner_join(avg.keqs.all, cele.all.df, by=c("primer", 'template', 'dG')) 
pred.cov.cele <- predict.coverage(cele.keqs.df, eps=eps.cele)  


```
```{r}
pred.cov.cele %>%
  ggplot(., aes(log(t.usage/sum(t.usage)), log(pred.cov/sum(pred.cov)), color=nuc)) + 
  geom_point(alpha=0.4) +
  geom_abline(slope=1, intercept=0, color='red') +
  theme_classic() +
  xlab("expected cov") + ylab("predicted cov") 

cor(pred.cov.cele$pred.cov, pred.cov.cele$t.usage)
```

Predicted coverage with BS-seq
```{r}
load("~/AvOwork/rdata/model_epsilon.RData")
epsilon.d3r <- exp(predict(model.epsilon, data.frame(genomecov=perc.d3R)))

d3r.keqs <- compute.keqs(d3r.all.df, eps=epsilon.d3r, filter.pt = 1)
pred.cov.d3r <- predict.coverage(d3r.keqs, eps = epsilon.d3r)
```


```{r}
cor(pred.cov.d3r$t.usage, pred.cov.d3r$pred.cov)
pred.cov.d3r %>%
  mutate(nuc=sapply(template, prevalent_nucleotide)) %>%
    ggplot(., aes(log(t.usage/sum(t.usage)), log(pred.cov/sum(pred.cov)), color=nuc)) + 
  geom_point(alpha=0.4) +
  geom_abline(slope=1, intercept=0, color='red') +
  theme_classic() +
  xlab("log(expected cov)") + ylab("log(predicted cov)") +
        theme(legend.title = element_blank(),
        legend.text = element_text(size=20),
        axis.text = element_text(size=20),
        axis.title = element_text(size=30),
        title = element_text(size=30)) +
  ggtitle("Mouse - BS") +
  ggsave("~/AvOwork/output/epsilon_estimate/predcov_BS_mm10_nofilt.pdf")

```

```{r, echo=F}
D3G <- load.pt.data("~/mnt/edann/primer_conc_VAN2493/D3G_tr2_R1_bismark_bt2.ptCounts.qualFilt.csv")
D3T <- load.pt.data("~/mnt/edann/primer_conc_VAN2493/D3T_tr2_R1_bismark_bt2.ptCounts.qualFilt.csv")
```
```{r}
mm10.abundance <- load.kmer.abundance("~/mnt/edann/hexamers/genomes_kmers/mm10.kmerAbundance.csv")
mfold.dg <- load.modelled.deltaG("~/mnt/edann/hexamers/rand_hex_deltaG_ions.txt.gz")
d3g.all.df <- join.pt.data(D3G$matches, D3G$t.usage, mm10.abundance, mfold.dg)
d3t.all.df <- join.pt.data(D3T$matches, D3T$t.usage, mm10.abundance, mfold.dg)

save(d3t.all.df, file='~/AvOwork/rdata/D3T_pt.RData')
save(d3g.all.df, file='~/AvOwork/rdata/D3G_pt.RData')
```

```{r}
perc.d3g <- 1-0.969021
epsilon.d3g <- exp(predict(model.epsilon, data.frame(genomecov=perc.d3g)))

d3g.keqs <- inner_join(select(d3r.keqs, primer, template, keq), d3g.all.df, by=c("primer", 'template')) 
prob.g <- batch.prob.uniform(nuc.probs = c(pA=0.25, pT=0.05, pC=0.25, pG=0.45))
pred.cov.d3g <- predict.coverage(d3g.keqs, eps = epsilon.d3g, prob = prob.g)

```

```{r}
pcc.d3g <- cor(pred.cov.d3g$t.usage, pred.cov.d3g$pred.cov)
pred.cov.d3g %>%
  mutate(nuc=sapply(template, prevalent_nucleotide)) %>%
    ggplot(., aes(log(t.usage/sum(t.usage)), log(pred.cov/sum(pred.cov)), color=nuc)) + 
  geom_point(alpha=0.4) +
  geom_abline(slope=1, intercept=0, color='red') +
  theme_classic() +
  xlab("log(expected cov)") + ylab("log(predicted cov)") +
        theme(legend.title = element_blank(),
        legend.text = element_text(size=20),
        axis.text = element_text(size=20),
        axis.title = element_text(size=30),
        title = element_text(size=30)) 

```

```{r}
perc.d3t <- 1-0.97141
epsilon.d3t <- exp(predict(model.epsilon, data.frame(genomecov=perc.d3t)))

d3t.keqs <- inner_join(select(d3r.keqs, primer, template, keq), d3t.all.df, by=c("primer", 'template')) 
prob.t <- batch.prob.uniform(nuc.probs = c(pA=0.25, pT=0.45, pC=0.25, pG=0.05))
pred.cov.d3t <- predict.coverage(d3t.keqs, eps = epsilon.d3t, prob = prob.t)

```

```{r}
plot.prediction <- function(pred.cov.df){
  pcc.pred <- cor(pred.cov.df$t.usage, pred.cov.df$pred.cov)
  pl <- pred.cov.df %>%
    mutate(nuc=sapply(template, prevalent_nucleotide)) %>%
      ggplot(., aes(log(t.usage/sum(t.usage)), log(pred.cov/sum(pred.cov)), color=nuc)) + 
    geom_point(alpha=0.4) +
    geom_abline(slope=1, intercept=0, color='red') +
    theme_classic() +
    xlab("log(observed cov)") + ylab("log(predicted cov)") +
          theme(legend.title = element_blank(),
          legend.text = element_text(size=20),
          axis.text = element_text(size=20),
          axis.title = element_text(size=30),
          title = element_text(size=30)) +
    ggtitle(paste('R.sq.', ' = ', round(pcc.pred,2), sep=''))
  return(pl)
  }

plot.prediction(pred.cov.d3g) + coord_cartesian(ylim=c(-22,-4)) + ggsave("~/AvOwork/output/primer_batch_VAN2493/obsVSexp_G_new.pdf")
plot.prediction(pred.cov.d3r) + coord_cartesian(ylim=c(-22,-4)) + ggsave("~/AvOwork/output/primer_batch_VAN2493/obsVSexp_R_new.pdf")
plot.prediction(pred.cov.d3t) + coord_cartesian(ylim=c(-22,-4)) + ggsave("~/AvOwork/output/primer_batch_VAN2493/obsVSexp_T_new.pdf")
```

When we switch probability vectors as a control we get very poor correlation, while usage and predicted coverage are stil somewhat correlated if we use as probability vector the even probability.
```{r}
pred.cov.wrong.d3t <- predict.coverage(d3t.keqs, eps = epsilon.d3t, prob = prob.g)
pred.cov.wrong.d3g <- predict.coverage(d3g.keqs, eps = epsilon.d3g, prob = prob.t)
pred.cov.wrong2.d3t <- predict.coverage(d3t.keqs, eps = epsilon.d3t)
pred.cov.wrong2.d3g <- predict.coverage(d3g.keqs, eps = epsilon.d3g)

pcc.d3g.wT <- cor(pred.cov.wrong.d3g$pred.cov, pred.cov.wrong.d3g$t.usage)
pcc.d3t.wG <- cor(pred.cov.wrong.d3t$pred.cov, pred.cov.wrong.d3t$t.usage)
pcc.d3g.wT; pcc.d3t.wG
```

This might be telling us something about the sensitivity of the method: probably the prediction cannot capture such fine grained differences in primer concentration.
```{r}

g.probs <- seq(0,0.5,0.05)
prob.vecs <- lapply(g.probs, function(x) batch.prob.uniform(nuc.probs = c(pA=0.25, pT=1-0.5-x, pG=x, pC=0.25)))
# prob.vec <- batch.prob.uniform(nuc.probs = c(pA=0.25, pT=0.05, pC=0.25, pG=0.45))
pc.d3t <- lapply(prob.vecs, predict.coverage, keqs.df=d3t.keqs, eps=epsilon.d3t)
pc.d3g <- lapply(prob.vecs, predict.coverage, keqs.df=d3g.keqs, eps=epsilon.d3g)
pc.d3r <- mclapply(prob.vecs, predict.coverage, keqs.df=d3r.keqs, eps=epsilon.d3r, mc.cores = 4)
```


```{r}
cor.pt.d3t <- sapply(pc.d3t, function(x) cor(x$pred.cov, x$t.usage))
cor.pt.d3g <- sapply(pc.d3g, function(x) cor(x$pred.cov, x$t.usage))
cor.pt.d3r <- sapply(pc.d3r, function(x) cor(x$pred.cov, x$t.usage))

data.frame(prob.G = g.probs, prob.T = 1-g.probs, 
           moreT = cor.pt.d3t, 
           moreG = cor.pt.d3g,
           random = cor.pt.d3r) %>%
  gather(key='sample', value='PCC', 3:5) %>%
  ggplot(., aes(prob.G, PCC, group=sample, color=sample)) +
  geom_line(size=1.5) +
  geom_point(size=2) +
  theme_minimal() +
  xlab('% G') + ylab(expression(R^2)) +
  scale_x_continuous(sec.axis=sec_axis(~0.5-., name='% T')) +
  scale_color_discrete(name='') +
  scale_color_brewer(palette = 'Accent') +
  theme(legend.title = element_blank(),
        legend.text = element_text(size=20),
        axis.text = element_text(size=20),
        axis.title = element_text(size=30),
        title = element_text(size=30)) +
  ggsave(filename = "~/AvOwork/output/primer_batch_VAN2493/variable_pcc.pdf",width = 12, height = 6)

```

```{r}
primer.usage.cele <- cele.all.df %>%
  group_by(primer) %>%
  summarise(p.usage= sum(pt)) %>%
  rename(hex=primer)
primer.usage.cele %>%
  mutate(p.usage.rank=dense_rank(p.usage)) %>%
  mutate(label=ifelse(p.usage.rank>784, hex, '')) %>%
  ggplot(., aes(p.usage.rank, p.usage)) + 
  geom_point() +
  # geom_bar(stat = 'identity') +
  geom_text_repel(aes(label=label))
  
```

### VAN2591 BS dataset

Load tables
```{r}
rdata.files <- list.files("~/mnt/edann/VAN2591/", pattern = 'Preamp-[^noBS].+.RDS', full.names = T)
pt.tabs.bs <- mclapply(rdata.files[grep(rdata.files, pattern = 'handMix')], readRDS, mc.cores = 4)
names(pt.tabs.bs) <- gsub(rdata.files[grep(rdata.files, pattern = 'handMix')], pattern = '.+/CG-BS-pbat-bulk-1xPreamp-|_lmerged_R1.ptCounts.qualFilt.csv.RDS', replacement = '')
str(pt.tabs.bs)
```

Compute epsilon
```{r}
gencov.all <- read.table("~/mnt/edann/VAN2591/genomecov_all.txt", sep='\t', header=F, col.names = c('smp', 'genomecov'))
gencov.BS <- gencov.all %>%
  mutate(genomecov=1-genomecov) %>%
  filter(!grepl(smp, pattern = 'noBS'))

epsilons <- exp(predict(model.epsilon, filter(gencov.BS, grepl(smp, pattern = 'handMix')) %>% select(genomecov)))
eps.bs <- filter(gencov.BS, grepl(smp, pattern = 'handMix')) %>%
  mutate(eps=epsilons)

keqs.tab.bs <- map(names(pt.tabs.bs), function(smp.name) compute.keqs(pt.tabs.bs[smp.name][[1]], eps = filter(eps.bs, smp==smp.name)$eps, filter.pt = 1))
names(keqs.tab.bs) <- names(pt.tabs.bs)



```

***** TO FINISH!!! *****

Predicted coverage computed with `run_cov_prediction_BS.r`

```{r}
pred.cov.files <- list.files("~/mnt/edann/VAN2591", pattern = 'BS.*.RData', full.names = T)

load(pred.cov.files[[1]])
cele.bs.pred.cov.batch <- output 

load(pred.cov.files[[2]], verbose = T)
drerio.bs.pred.cov.batch <- output 

load(pred.cov.files[[4]])
mm.bs.pred.cov.batch <- output 

```
```{r}
map(seq_along(cele.pred.cov.batch$pred.coverage), function(i) plot.prediction(cele.pred.cov.batch$pred.coverage[[i]], main=names(cele.pred.cov.batch$pred.coverage)[i]))
```

```{r}
map(seq_along(drerio.pred.cov.batch$pred.coverage), function(i) plot.prediction(drerio.pred.cov.batch$pred.coverage[[i]], main=names(drerio.pred.cov.batch$pred.coverage)[i]))
```

```{r}
# map(seq_along(human.pred.cov.batch$pred.coverage), function(i) plot.prediction(human.pred.cov.batch$pred.coverage[[i]], main=names(human.pred.cov.batch$pred.coverage)[i]))
```

```{r}
map(seq_along(mm.pred.cov.batch$pred.coverage), function(i) plot.prediction(mm.pred.cov.batch$pred.coverage[[i]], main=names(mm.pred.cov.batch$pred.coverage)[i]))

```

```{r}
plot.batch.accuracy(mm.bs.pred.cov.batch$pcc.primer.batch)
# plot.batch.accuracy(human.pred.cov.batch$pcc.primer.batch)
plot.batch.accuracy(cele.bs.pred.cov.batch$pcc.primer.batch)
plot.batch.accuracy(drerio.bs.pred.cov.batch$pcc.primer.batch)
```

